# (APPENDIX) Appendix {-} 

#Mortality Population Risk Tool (MPoRT) 

**Outcomes: 5-yr risk of death, Life Expectancy, Cause-deleted Life Expectancy**

**Calculations**
 
Using MPoRT you are able to calculate:

- 5 year mortality risk
- Number of deaths
- Life Expectancy
- Cause-deleted deaths and life expectancy
- Burden of health behaviour in deaths and on life expectancy

**Types of Questions**

- What is the burden of smoking on life expectancy?
- How many deaths would be prevented if everyone met their daily excercise requirements?



**Description**: A multivariable predictive risk model that estimates the future risk of all-cause death in Canada. It adjusts for health behaviours: smoking, unhealthy alcohol consumption, poor diet, and physical inactivity, and a wide range of other risk factors. 

Versions of MPoRT have been developed since 2012 and used in various studies. Each version of MPoRT (v1.0, v1.2, v2.0) used the Ontario subset of the Canadian Community Health Survey (CCHS) for development and the survey respondents were linked to personal death records. In later versions of MPoRT (v1.2, v2.0) the following changes were made:, (a) algorithm variables were adjusted to improve predictions, and (b) the algorithms were validated using: the Ontario subset of CCHS of the years that were not used in development and the National CCHS dataset (excluding Ontario).


**MPoRTv1.0**
Was used in the "Seven More Years" report, a joint report with Public Health Ontario and IC/ES (https://www.ices.on.ca/Publications/Atlases-and-Reports/2012/Seven-More-Years). 
In summary, the algorithm estimated the risk of death associated with health behaviours: smoking, unhealthy alcohol consumption, poor diet, physical inactivity and stress. There were approximately 550,000 person-years of follow up and over 6000 deaths in the development dataset. The algorthim used categorical predictor variables for health behaviours and sociodemographic factors.

**MPoRTv1.2**
Was published in PLoS (https://journals.plos.org/plosmedicine/article?id=10.1371/journal.pmed.1002082).
In summary, the algorithm estimated the risk of death associated with health behaviours: smoking, unhealthy alcohol consumption, poor diet, and physical inactivity (stress was removed due to its low prediction ability).  There were approximately 1 million person-years of follow up and over 9000 deaths in the development and validation datasets. The algorithm used multiple continous predictor variables, and added chronic disease predictor variables and interaction terms.

**MPoRTv2.0 - The version used in Project Big Life's Planning Tool**
This version of MPoRT has not yet been published.

*Development*: This predictive risk model was developed using Ontario subsets of the 2001 to 2008 CCHS and participants were linked to personal health records. There were approximately 1.3 million person-years of follow-up and over 15,000 deaths in the developmental dataset.

*Validation*: This predictive risk model was validated using three different datasets: Ontario subset of the 2009 to 2012 CCHS, National dataset (except Ontario) of the 2003 to 2008 CCHS, and the National dataset of the 2000 and 2005 National Health Interview Survey in the United States of America. In all validation datasets individuals were linked to personal health records.

*Parameters*: The parameters used in this predictive risk model are: 
```{r, echo=FALSE, results='hide', message=FALSE}
library(knitr)
library(kableExtra)
#Replicated Table 1 from MPoRTv2 Figures - Predictor variables for the MPoRTv2
(pred.MPoRT2 <- data.frame(
    Category = c(rep("Demographic",2), rep("Health Behaviour",9), rep("Socio-demographic",15), rep("Chronic Conditions",10)),
    Variable = c("Age*", "Sex", "Pack years of smoking", rep("Smoking Status",4),
               "Alcohol (number of drinks per week)", "Former/non-drinker", "Simplified diet score",
               "Leisure physical activity (MET)", rep("Ethnicity",6),"Immigrant",
               "Fraction of lifetime in Canada", rep("Education",4), rep("Neighbourhood social and material
                                                                         deprivation",3), 
               "Diabetes", "High Blood Pressure", "Chronic Respiratory Disease", "Mood Disorder", "Cancer",
               "Dementia", "Heart Disease", "Stroke", "Epilepsy", "BMI"), 
       Scale = c("Continous", "Dichotomous", "Continous", "Categorical", " ", " ", " ", "Continous",
                 "Dichotomous", "Continous", "Continous", "Categorical", " ", " "," "," "," ",
                 "Dichotomous", "Continous", "Categorical"," "," "," ","Ordinal"," "," ", "Dichotomous",
                 "Dichotomous", "Dichotomous","Dichotomous", "Dichotomous", "Dichotomous", "Dichotomous",
                 "Dichotomous", "Dichotomous",
                 "Continous"), 
    Description = c("5 knot spline. Valid range 20 to 102", "Stratified Female/Male", 
                    "3 knot spline. Valid range: 0 to 78 (Female), 0 to 112.5 (Male)", "Non-smoker", 
                   "Current Smoker", "Former Smoker <= 5 years", "Former > 5 years", 
                   "4 knot spline (Females) and 3 knot spline (Males). Valid range: 0 to 25 (Female), 0 to 50 (Male)", 
                   "Yes/No",
                   "3 knot spline. Valid range: -18.9 to 20.7 (Female), -16.8 to 18.4 (Male)", 
                   "3 knot spline. Valid range: 0 to 12.4 (Female), 0 to 16 (Male)",
                   "White", "Black", "Chinese", "Arab; South Asian; West Asian",
                   "Filipino; Japanese; Korean; Southeach Asian",
                   "Other; Indigenous; Latin American; Multiple origin; unknown", "Yes/No", 
                   "3 knot spline^†^. Valid range: 0 to 1", "Less than secondary", 
                   "Secondary School Graduation", "Some Post-Secondary", "Post-Secondary Graduation",
                   "Low (1st or 2nd quantile", "High (4th or 5th quantile)", "Moderate (all others)",
                   "Yes/No", "Yes/No","Yes/No", "Yes/No", "Yes/No", "Yes/No", "Yes/No", "Yes/No",
                   "Yes/No^‡^", "3 knot spline. Valid range: 8.9 to 47.2 (Female), 8.6 to 43.7 (Male)"))   )#closingbracket 

#Adding in symbols for footnotes, *Still unsure how to put a footnote in a cell of the table and not only the column or row name:
#footnote_pred.MPoRT2 <- pred.MPoRT2
#names(footnote_pred.MPoRT2)[2] <- paste0(names(footnote_pred.MPoRT2)[2], footnote_marker_symbol(1, "html"))
```
```{r, echo=FALSE}
#Creating the table
kable(pred.MPoRT2) %>%
  kable_styling(full_width = T) %>%
  collapse_rows(columns = 1:2, valign = "top") %>%
  scroll_box(height = "400px") %>%
#Adding the footnote 
  footnote(symbol = c("Age interaction included for all variables exept immigrant, fraction of time in Canada, and ethnicity", "Excluded in the male model, remains in the female model", "Excluded in the female model, remains in the male model"), footnote_as_chunk = TRUE)
```
