--- 
title: "Planning Tool Guidance"
author: "Project Big Life"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [./bibliography/book.bib, ./packages.bib]
biblio-style: apalike
link-citations: yes
always_allow_html: yes
github-repo: Big-Life-Lab/BATCH-Guidance
description: "Guidance on how to use the Project Big Life Planning Tool."
---

# Welcome to Project Big Life's Planning Tool

```{r eval=FALSE, echo=FALSE}
library(bookdown)
library(knitr)
library(kableExtra)
install.packages("ggplot2")
#install.packages('tinytex')
#tinytex::install_tinytex()
tinytex:::is_tinytex()
```

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

_TO DO: Insert the image for the PBL planning tool_

###What is the Project Big Life Planning Tool

The Project Big Life Planning Tool supports public health professionals research, plan, develop, and evaluate evidence-based health interventions in just a few steps. 

The tool can help you:

- Determine who in the population is at risk
- Predict number of new cases of the disease for a population
- Determine how much a risk factor contributes to a health outcome
- Assess health benefits of potential interventions

The tool is accessible to everyone and easy to use.


###Who made the Project Big Life Planning Tool?

The Project Big Life Planning Tool was developed by the [Project Big Life](https://www.projectbiglife.ca) Team based at the [Ottawa Hospital](http://www.ohri.ca/home.asp) and the [Bruyère Research Institute](https://www.bruyere.org/en/bruyere-research-institute).

The planning tool was made in collaboration with [Public Health Ontario](https://www.publichealthontario.ca/), the [Population Health Analytics Lab](https://pophealthanalytics.com/) at the University of Toronto and [Ottawa Public Health](http://www.ottawapublichealth.ca/en/index.aspx). 

The tool supports the [Seven More Years](https://www.ices.on.ca/Publications/Atlases-and-Reports/2012/Seven-More-Years) report, a study that examined the health behaviour attributale mortality in Ontario. Algorithms used in the planning tool were developed using linked population data housed at [ICES](https://www.ices.on.ca/) and validated at Statistics Canada. All algorithms have been published in peer-review journals. See [Project Big Life](https://www.projectbiglife.ca/science) or [appendices](#mport) to the specific algorithms.

