--- 
title: "Project Biglife Planning Tool Guide"
author: "Project Big Life"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [./bibliography/book.bib, ./packages.bib]
biblio-style: apalike
link-citations: yes
always_allow_html: yes
github-repo: Big-Life-Lab/BATCH-Guidance
description: "Guidance on how to use the Project Big Life Planning Tool."
---

# Introduction

```{r eval=FALSE, echo=FALSE}
library(bookdown)
library(knitr)
library(kableExtra)
install.packages("ggplot2")
#install.packages('tinytex')
#tinytex::install_tinytex()
tinytex:::is_tinytex()
```

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

The Project Big Life Planning Tool ([PBL Planning Tool](http://planning.projectbiglife.ca/)) is a web-based application to assess the burden of health behaviours (smoking, diet, physical activity) and the potential health impact of public health preventive strategies in Canada. 

The tool was developed in response to requests from public health departments to create, replicate, and update health status reports such as Public Health Ontario's [Seven More Years](https://www.ices.on.ca/Publications/Atlases-and-Reports/2012/Seven-More-Years) which examined the impact of health behaviours on health outcomes: mortality and life expectancy.

We hope you find the tool helpful for your research and policy planning, including the development and evaluation of population-based health interventions.

**The tool can help you to:**

- Determine who in the Canadian population is at risk of dying
- Predict the number of new deaths in different Canadian populations
- Determine how much a risk factor contributes to a health outcome
- Assess the health benefits of potential interventions

**Examples questions the tool can answer**

The Project Big Life Planning Tool can answer the following types of questions:

- What is the burden of smoking on life expectancy in Canada?
- How many deaths would be prevented if everyone in Winnipeg increased their daily exercise by 10%?

The tool is well-suited to calculate health burdens for a range of social and demographic groups, providing a resource to examine inequities associated with health behaviours.[@manuel2018] Also supported is the ability to examine the impact of different types of preventive scenarios including population strategies that have a small incremental change in health behaviours for a broad target population or high-risk strategies with a larger change in health behaviours that target a smaller number of high-risk people.[@PoRTover]

**How it works**

The PBL planning tool uses multivariable predictive risk algorithms. All algorithms use by the planning too are developed and validated using data routinely collected by Statistics Canada and provincial health agencies, and the algorithms have been published in various journals. More information about algorithms can be found in the [Background](#background), [Key concepts](#keyconcepts) and [Appendices](#mport).

**Why should I use the tool**

- It is **easy** and **flexible** to use. 

    - The user only needs to upload their data and choose which calculation to run.
    - It can be used to assess the future risk of a health outcome.
    - It can be used to assess the effectiveness of different intervention scenarios (e.g. policy) on a health outcome.

- It is **robust** and **flexible**.

    - It can be used to examine health outcomes in a range of social and demographic groups. 
    
-    It generates **accurate** predictions.

    - The tool's algorithms have been published in peer-review journals, validated in the Canadian population for a wide range of populations, and are publicly available.
    
-    It is **private**. 

    - Loaded data remains on your computer and is not uploaded or sent anywhere.