--- 
title: "Planning Tool Guidance"
author: "Project Big Life"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
always_allow_html: yes
github-repo: Big-Life-Lab/BATCH-Guidance
description: "Guidance on how to use the Project Big Life Planning Tool."
---

# Welcome to Project Big Life's Planning Tool

```{r eval=FALSE, echo=FALSE}
library(bookdown)
library(knitr)
library(kableExtra)
install.packages("ggplot2")
#install.packages('tinytex')
#tinytex::install_tinytex()
tinytex:::is_tinytex()
```

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

_TO DO: Insert the image for the PBL planning tool_

###What is the Project Big Life Planning Tool

*To do: Develop video showcasing the platform including what it is and why someone should use the platform*

###Who made the Project Big Life Planning Tool?

The Project Big Life Planning Tool was developed by the Project Big Life Team. The Project Big Life Team is part of the ICES. The below video explains what ICES is. 

```{r embed2, out.extra='style="border: none;"', out.width='99%', echo=FALSE}
knitr::include_url('https://www.youtube.com/embed/5pZhRSM1cyk')
#Not sure how to work the vimeo embeding
#knitr::include_url('https://vimeo.com/embed/267453408')
```